Solving an optimization problem employing restraint programming

### 2015-12-13 proj2/doc/article.tex
source code in the annex

>>>>

\section{Annex}

\subsection{Source code}

\begin{verbatim}
:-use_module(library(clpfd)).
:-use_module(library(aggregate)).
:-use_module(library(lists)).

clearScreen:-write('\e[2J').

professoresQueDaoLingua(_, [], []).
professoresQueDaoLingua(Lingua,  [ [Prof, Linguas,_] | OutrosProfs], Lista):-
	member(Lingua-_, Linguas),
	professoresQueDaoLingua(Lingua, OutrosProfs, Lista2),
	append([Prof], Lista2, Lista).

professoresQueDaoLingua(Lingua,  [ [_, _,_] | OutrosProfs], Lista):-
	professoresQueDaoLingua(Lingua, OutrosProfs, Lista).

generate_professor_availability([], []).
generate_professor_availability([[Indice,_,Dias]|Professors], Table):-
	generate_professor_availability(Dias,Indice,TableThis),
	generate_professor_availability(Professors, TableRest),
	append(TableThis, TableRest, TableN),
	append([[-1,0],[-1,1],[-1,2],[-1,3],[-1,4],[-1,5]], TableN, Table).

generate_professor_availability([], _, []).
generate_professor_availability([Dia|Dias],Indice, Table):-
	generate_professor_availability(Dias, Indice, TableRest),
	append([[Indice,Dia]], TableRest, Table).

initialize_days_list([],_,_,_,_).
initialize_days_list([H|T], N, NProfessores, FdSetDeProfs, ProfAvailable):-
	H = [N,Hora,Prof],
	Prof #= -1 #<=> Hora #= 0,
	findall(ProfA, member( [ProfA, N], ProfAvailable), ListaDeProfsA),
	list_to_fdset(ListaDeProfsA, FdSetDisponivel),
	fdset_intersection(FdSetDisponivel, FdSetDeProfs, FdSetFinal),
	Prof in_set FdSetFinal,
	Hora in {0} \/{2},
	N1 is N+1,
	initialize_days_list(T, N1, NProfessores, FdSetDeProfs, ProfAvailable).

initialize_solution([], _,_,_,_).
initialize_solution([H|T], Lingua, NProfessores, Professores, ProfAvailable):-
	H = Lingua-Lista,
	length(Lista, 5),
	professoresQueDaoLingua(Lingua, Professores, ListaDeProfs),
	list_to_fdset([-1|ListaDeProfs], FdSetDeProfs),
	initialize_days_list(Lista, 0, NProfessores,FdSetDeProfs,ProfAvailable),
	Lingua1 is Lingua+1,
	initialize_solution(T, Lingua1, NProfessores, Professores, ProfAvailable).

	restrict_course_times([]).
	restrict_course_times([_-[[_,H1,_],[_,H2,_],[_,H3,_],[_,H4,_],[_,H5,_]]|T]):-
		sum([H1,H2,H3,H4,H5],#=<,4),
		sum([H1,H2,H3,H4,H5],#\=,1),
		sum([H1,H2,H3,H4,H5],#>=,0),
		restrict_course_times(T).

labelAll(H,P,Cost):- append(H, P, List), labeling([up,bisect,ff,timeout(60000,),
maximize(Cost)], List).

restrict_teacher_hours(_,_,[], _).
restrict_teacher_hours(Horas, ProfessoresDominio, [[Indice,_,_]|T], NProfs):-
	make_prof_table(Indice, Table, NProfs),!,
	append(Table, [[-1, 0]], NTable),
	restrict_teacher(NTable, ProfessoresDominio, Rest),
	sum_product(Horas, Rest, Sum),
	Sum #=< 15,
	restrict_teacher_hours(Horas, ProfessoresDominio, T, NProfs).

restrict_teacher(_, [], []).
restrict_teacher(Table, [P|Ps], [Rest|T]):-
	table([[P,Rest]], Table),
	restrict_teacher(Table, Ps, T).

sum_product([],[],0).
sum_product([X|Xs],[Y|Ys], Sum):-
	sum_product(Xs,Ys,SumRest),
	Sum #= X*Y + SumRest.

make_prof_table(Indice, Lista, NProfs):-
	Length is NProfs+1,
	length(Lista, Length),
	iterate_prof_table(Lista, NProfs, Indice).

iterate_prof_table(_,-1,_).
iterate_prof_table(Lista, NProfs, NProfs):- !, nth0(NProfs, Lista, [NProfs, 1]),
 Next is NProfs-1, iterate_prof_table(Lista, Next, NProfs).
iterate_prof_table(Lista, NProfs, Indice):- nth0(NProfs, Lista, [NProfs,0]), 
Next is NProfs-1, iterate_prof_table(Lista, Next, Indice).

generate_domain_day_list([],[],[]).
generate_domain_day_list([[_,Hora,Prof]|Dias], H, P):-
	generate_domain_day_list(Dias, HRest, PRest),
	append([Hora], HRest, H),
	append([Prof], PRest, P).

generate_variable_domain_list([],[],[]).
generate_variable_domain_list([_-Dias|Solucoes], H, P):-
	generate_domain_day_list(Dias, HThis, PThis),
	generate_variable_domain_list(Solucoes, HRest, PRest),
	append(HThis, HRest, H),
	append(PThis, PRest, P).

custo_nulo([], 0).
custo_nulo(TabelaCustos, NLinguas):-
	NLinguas \= 0,
	N is NLinguas-1,
	custo_nulo(TabelaCustosR, N),
	append([[-1,N,0]], TabelaCustosR, TabelaCustos).

escola_de_linguas(Professores, Candidaturas, Linguas, Rooms, Lucro, Solucao):-
	length(Professores, NProfessores),
	length(Linguas, NLinguas),
	length(Solucao, NLinguas),
	generate_professor_availability(Professores, ProfAvailable),
	cria_tabela_custos(Professores, TabelaCustosReais, []),
	custo_nulo(CustoNulo, NLinguas),
	append(CustoNulo, TabelaCustosReais, TabelaCustos),
	initialize_solution(Solucao, 0, NProfessores, Professores,ProfAvailable),
	generate_days_prof_table(Solucao, DayProf),
	restrict_rooms(DayProf,Rooms, NLinguas),
	generate_variable_domain_list(Solucao, H, P),
	restrict_teacher_hours(H, P, Professores, NProfessores),
	restrict_course_times(Solucao),
	make_profit(Solucao, Linguas, Professores, TabelaCustos, Candidaturas, Lucro),
	labelAll(H, P, Lucro),
	printSolution(Solucao),
	 write('O lucro total obtido e de '), write(Lucro),  nl,
	fd_statistics.

cria_tabela_custos([], TabelaCustos, TabelaCustos).
cria_tabela_custos([Professor|Professores], TabelaCustos, Acum):-
	Professor = [Indice,Linguas,_],
	cria_tabela_linguas(Linguas, Indice, TabelaLinguas, []),
	append(Acum, TabelaLinguas, NovaTabela),
	cria_tabela_custos(Professores, TabelaCustos, NovaTabela).

cria_tabela_linguas([], _, T, T).
cria_tabela_linguas([LinguaCusto|LinguasCustos], Indice, TabelaLinguas, Acum):-
	LinguaCusto = IndiceL-Custo,
	append(Acum, [[Indice,IndiceL,Custo]], NovaTabela),
	cria_tabela_linguas(LinguasCustos, Indice, TabelaLinguas, NovaTabela).

make_profit([], _, _, _,_, 0).
make_profit([Solucao|Solucoes], Linguas, Professores,TabelaCustos, Candidaturas, Lucro):-
	make_profit_language(Solucao, Linguas, Professores, TabelaCustos, Candidaturas
	, LucroLingua),
	make_profit(Solucoes, Linguas, Professores, TabelaCustos, Candidaturas, LucroResto),
	Lucro #= LucroResto+LucroLingua.

	Inscritos is min(NumeroCandidatos, 15),
	Lucro #= Preco*Inscritos*Horas,

	LucroDiaAtual #= (Lucro-Prejuizo),
	make_profit_day(OutrosDias, Linguas, Professores,TabelaCustos, Candidaturas, IndiceL
	, LucroRest),
	LucroDia #= LucroRest + LucroDiaAtual.

generate_days_prof_table([], [[],[],[],[],[]]).
generate_days_prof_table([_-[[_,_,Prof1],[_,_,Prof2],[_,_,Prof3],[_,_,Prof4],
[_,_,Prof5]]|Solucoes], [Dia1, Dia2, Dia3, Dia4, Dia5]):-
	generate_days_prof_table(Solucoes, [Dia1R, Dia2R, Dia3R, Dia4R, Dia5R]),
	append([Prof1], Dia1R, Dia1),
	append([Prof2], Dia2R, Dia2),
	append([Prof3], Dia3R, Dia3),
	append([Prof4], Dia4R, Dia4),
	append([Prof5], Dia5R, Dia5).

restrict_rooms([],_,_).
restrict_rooms([Dia|Dias], Rooms, NLinguas):-
	AulasVazias is NLinguas - Rooms,
	count(-1, Dia, #>=, AulasVazias),
	restrict_rooms(Dias, Rooms, NLinguas).

printSolution([]).
printSolution([Indice-Dias|Solutions]):-
	nl, write('Language '), write(Indice), write(' Time Table'),nl, write('_______________'),
	nl,
	printLanguage(Indice, Dias),
	printSolution(Solutions).

dia(0, 'Monday').
dia(1, 'Tuesday').
dia(2, 'Wednesday').
dia(3, 'Thursday').
dia(4, 'Friday').

printLanguage(_, []).
printLanguage(_, [[_,0,-1]|Dias]):-printLanguage(Indice, Dias).
printLanguage(Indice, [ [Dia,Horas,Prof]|Dias]):-
	dia(Dia, DiaDaSemana),
	write(DiaDaSemana), write(' - '), write(Horas), write(' hours with professor '), 
	write(Prof), nl,
	printLanguage(Indice, Dias).

\end{verbatim}
